---
title: 'DATA 607 Assignment #3'
author: "Kristin Lussi"
output:
  html_document: default
  pdf_document: default
date: "2023-09-24"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1:

Using the 173 majors listed in fivethirtyeight.com's College Majors
dataset
[<https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/>],
provide code that identifies the majors that contain either "DATA" or
"STATISTICS"

```{r, include = FALSE, warning = FALSE}
library(readr)
library(stringr)
library(dplyr)
```

***Answer:***

```{r, warning = FALSE}
# retrieve the csv file from GitHub
urlfile = "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"

collegemajors <- read_csv(url(urlfile), show_col_types = FALSE)

# filter the Major column of the collegemajors database to include only strings that contain the words "data" or "statistics"
data_or_stats <- collegemajors %>%
  filter(str_detect(Major, "DATA | STATISTICS")) 

# select only the major column to show the filtered majors
filtered_majors <- data_or_stats %>%
  select(Major) %>%
  distinct(Major)

# change the tibble to a data frame
filtered_majors <- data.frame(filtered_majors)

# show the data frame
filtered_majors
```

## Question 2:

Write code that transforms the data below:

[1] "bell pepper" "bilberry" "blackberry" "blood orange" [5] "blueberry"
"cantaloupe" "chili pepper" "cloudberry"\
[9] "elderberry" "lime" "lychee" "mulberry"\
[13] "olive" "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry",
"cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime",
"lychee", "mulberry", "olive", "salal berry")

***Answer:***

```{r}
library(stringr)

# recreate the character vector
fruit_vector <- c(
  "bell pepper", "bilberry", "blackberry", "blood orange",
  "blueberry", "cantaloupe", "chili pepper", "cloudberry",
  "elderberry", "lime", "lychee", "mulberry",
  "olive", "salal berry"
)

# convert the character vector into a string
fruit_string <- str_c("c(", str_c('"', fruit_vector, '"', collapse = ", "), ")", sep = "")

# Remove backslashes from the fruit string
formatted_string <- gsub("\\\\", "", fruit_string, fixed = TRUE)

# Print the fruit string
cat(formatted_string, sep = "\n")
```

## Question 3:

Describe, in words, what these expressions will match: a. (.)\\1\\1 b.
"(.)(.)\\2\\1" c. (..)\\1 d. "(.).\\1.\\1" e. "(.)(.)(.).\*\\3\\2\\1"

***Answer:***

a.  (.)\\1\\1 will match any 3 consecutive characters in a string. For
    example, "aaa".

b.  "(.)(.)\\2\\1" will match the a string where the first and second
    characters are repeated in reverse order. For example, "abba",
    "1221", etc.

c.  (..)\\1 captures 2 consecutive characters and stores them in the
    same capturing group. This will match any expression that have a
    repeated pair of characters (i.e. banana)

d.  "(.).\\1.\\1" captures 3 consecutive characters from a string. The
    first character is captured by the group, the second character is
    matched by the ".", and the third character is matched by the final
    ".". This will match any expression that has similar 3-character
    sequences, i.e. "axaxa", "cbcbc", or "12121".

e.  "(.)(.)(.).*\\3\\2\\1" captures 3 characters into 3 separate
    capturing groups, then the ".*" outside the "()" matches 0 or more
    characters. "\\3\\2\\1" matches the same characters as previously
    captured, but in reverse order. Examples: "xyz123zyx", "123456321",
    or "hannah".

## Question 4:

Construct regular expressions to match words that: 

- Start and end with
the same character. 

- Contain a repeated pair of letters (e.g. "church"
contains "ch" repeated twice.) 

- Contain one letter repeated in at least
three places (e.g. "eleven" contains three "e"s.)

***Answer:***

Start and end with the same character: "(.).\*\\1" Example:

```{r}
stringx <- "abcba"
str_view(stringx, "(.).*\\1")
```

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated
twice): "(..).\*\\1"

```{r}
church <- "church"
str_view(church, "(..).*\\1")
```

Contain one letter repeated in at least three places (e.g. "eleven"
contains three "e"s.): "

```{r}
eleven <- "eleven"
str_view(eleven, "([a-z]).*\\1.*\\1")
```

## Sources

Title: "The Economic Guide to Picking a College Major."
Author/Organization: FiveThirtyEight. URL:
<https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/>

Title: GitHub Repository for FiveThirtyEight College Majors Data
Author/Organization: FiveThirtyEight URL:
<https://github.com/fivethirtyeight/data/blob/15f210532b2a642e85738ddefa7a2945d47e2585/college-majors/majors-list.csv>
